---
title: "Gov 50 Final Project"
author: "Olanrewaju Adeyanju"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---



## Project thoughts

Are governments in Africa more likely to use repression or violence against events (violent or nonviolent actions taken) that are located in urban areas? In this study, I plan to examine the relationship between the type of location of events and if they are faced with violence or repression from the government. I hypothesize that events taken at capital cities or major urban areas are faced with repression by the government, and events that are taken at more rural locations face relatively lower repression. I expect to find that events taken at capital cities and major urban areas face more repression. This is because major urban cities allow for easier mobilization and resource allocation for events (violent or nonviolent actions taken) and therefore pose a bigger threat to the government and are consequently faced with more repression. My data set consists of violent and nonviolent actions taken in Africa. My unit of analysis is an event (violent or nonviolent actions taken). My explanatory variable of interest is the location of events or action taken. (Coding of the event locality 1 = Capital city; 2 = Other major urban area (population greater than 100,000); 3 = Rural (including small towns, villages with population less than 100,000); 4 = Multiple urban areas; 5 = Multiple rural areas; 6 = Province/region listed, exact location unknown; 7 = Nationwide. Effects several cities and rural areas;  -99 = location unknown). My outcome variable of interest is "repress", whether the government used repression or violence against participants in the event. Coding of the repression variable 0 = no repression used, 1 = non-lethal repression used (e.g. tear gas, arrests, etc), 2 = lethal repression used (there must be deaths reported). This variable is measured from the Social Conflict Analysis Database (SCAD). If I observe greater repression usage for actions in major cities or urban areas, this would provide support for my hypothesis. If, on the other hand, I observe lower repression usage or the same level of repression as in rural areas, this would provide evidence against my hypothesis. When I run my regression of the count of events in urban areas and major cities on the repression variable, a positive, significant coefficient would indicate support for my hypothesis.


## Milestone 3: Data Visualization


```{r}
library(tidyverse)

Data1 <- read_csv("C:\\Users\\temit\\OneDrive\\GOV50\\Final GOV 50 Project\\SCAD2018Africa_Final.csv")


# Create a binary variable for urban or rural
Data1 <- Data1 |>
  mutate(urban_rural = case_when(
    locnum %in% c(1, 2, 4) ~ "Urban",
    locnum %in% c(3, 5) ~ "Rural",
    TRUE ~ "Other" # For the remaining locations
  ))


# Rename the "repress" variable to "GovResponse"
Data1 <- Data1 |>
  mutate(GovResponse = factor(repress, levels = c("0", "1", "2")))

# Create a data visualization
Data1 |>
  group_by(urban_rural, GovResponse) |>
  summarize(Count = n()) |>
  group_by(urban_rural) |>
  mutate(Proportion = Count / sum(Count)) |>
  ggplot(aes(x = urban_rural, y = Proportion, fill = GovResponse)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Government Response by Event Location",
    x = "Event Location",
    y = "% of Repression Tactics",
    fill = "Government Response"
  ) +
  scale_fill_manual(
    values = c("0" = "steelblue", "1" = "yellow", "2" = "indianred"),
    labels = c("No Repression Used", "Non-lethal Repression Used", "Lethal Repression Used")
  ) +
  theme_minimal()


```


## Milestone 4: Analysis

This second visualization analyzes the outcome variable, 'Government response', in relation to another categorical variable, 'issue1'. The focus is on understanding the variation in government use of repression during events, based on the first issue mentioned as the primary source of tension leading to the taken action (violent or nonviolent). The initial visualization in milestone 3 did not provide a substantial difference in government repression across our x variable, location. To better understand the factors influencing government repression, we utilize the 'issue1' categorical variable, testing it across each category (1 = elections; 2 = economy, jobs; 3 = food, water, subsistence; 4 = environmental degradation; 5 = ethnic discrimination, ethnic issues; 6 = religious discrimination, religious issues; 7 = education; 8 = foreign affairs/relations; 9 = domestic war, violence, terrorism; 10 = human rights, democracy; 11 = pro-government; 12 = economic resources/assets; 13 = other; 14 = unknown, not-specified). This analysis reveals significant variation in the types of issues prompting government repression.


```{r}

# Create a second data visualization showing how the type of issue influences repression tactic
Data2 <- subset(Data1, issue1 != -99)

# Name my issue1 categorical variable
Data2 <- Data2 |>
  mutate(issue1 = case_when(
    issue1 == 1 ~ "elections",
    issue1 == 2 ~ "economy, jobs",
    issue1 == 3 ~ "food, water, subsistence",
    issue1 == 4 ~ "environmental degradation",
    issue1 == 5 ~ "ethnic discrimination, ethnic issues",
    issue1 == 6 ~ "religious discrimination, religious issues",
    issue1 == 7 ~ "education",
    issue1 == 8 ~ "foreign affairs/relations",
    issue1 == 9 ~ "domestic war, violence, terrorism",
    issue1 == 10 ~ "human rights, democracy",
    issue1 == 11 ~ "pro-government",
    issue1 == 12 ~ "economic resources/assets",
    issue1 == 13 ~ "other",
    issue1 == 14 ~ "unknown, not-specified"
  ))


Data2 |>
  group_by(issue1, GovResponse) |>
  summarize(Count = n()) |>
  group_by(issue1) |>
  mutate(Proportion = Count / sum(Count)) |>
  ggplot(aes(x = issue1, y = Proportion, fill = GovResponse)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Government Response by Issue Leading to Action",
    x = "First issue that started action taken",
    y = "% of Repression Tactics",
    fill = "Government Response"
  ) +
  scale_fill_manual(
    values = c("0" = "steelblue", "1" = "yellow", "2" = "indianred"),
    labels = c("No Repression Used", "Non-lethal Repression Used", "Lethal Repression Used")
  ) +
  theme(axis.text.x = element_text(angle = 90)) +
  theme_minimal()


```


